Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-09-01 FST-TCS/rebuttal/rebuttal.md

I enjoyed the running example of automatic music transcription which serves to illustrate every introduced concept and brings a different domain into consideration.  The overall impression of the paper is spoiled by many careless mistakes, such as formulas sticking out into the page margin (see below for a list).

Overall, I think that the paper cannot be published in its present form because the introduction of symbolic weighted automata without reference to the original literature could be seen as plagiarism.  I think, however, that the original work presented in the paper can be of interest to the scientific community.  This is why I opt to weakly reject the paper.  In order for the paper to be published, its focus should be shifted from the theory of symbolic weighted automata to the framework of symbolic weighted parsing.

In particular, the authors could spend fewer pages repeating already established
concepts from the literature and integrate Appendix D into the main paper.

REMAINS TO DO :
references :
- [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
- [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
Moreover, the `corresponds' is wrong, since superiority generalizes non-negative edge weights
- [ ] PS: Knuth's definition of superiority includes monotonicity
- [ ] 143: Lemma 4 is Lemma 3 in [21]
- [ ] 195: references: there is earlier research on symbolic automata, e.g. Veanes (2013)
- [ ]  Ref [22] and [23] both seem to refer to the same work by Mohri

- [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
- [ ] 57: weighted automata don't have ASTs, but runs
- [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
- [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
- [ ] 136: what implies?
- [ ] 153: quantify $x$ and the family ($y_i$) for $i ∈ ℕ$
- [ ] 155--156: alphabets are usually finite and non-empty sets
- [ ] 159: denoted by; also it seems like $\bar\Phi$ is a ($\Sigma \cup \Sigma \times \Delta$)-indexed family, why don't you introduce it like this
- [ ] 161--171: use an itemize environment 
- [ ] 162--165: I think these should also hold for $\Delta$, why don't you write this explicitly?
- [ ] 188: the definition of $\bigoplus_\Delta$ can be seen to be implicit.  Please fully quantify the alphabets of a label theory
- [ ] 208: What does $T$ compute?
- [ ] 219: resp.\@
- [ ] 220--221: you mean that a term CONTAINING $w_{ij}(…)$ equal to 0 is ignored
- [ ] 238: text in the margin, also it is not clear whether the if refers to the entire line or just the right definition
- [ ] 284: $x$ is a variable, so it should be typeset in math mode
- [ ] 291: I'd at least say once that $\Delta_i$, $\Delta_c$, and $\Delta_r$ are pairwise disjoint (or explain the symbol)
- [ ] 297: the extension here is not as in Section 3, there, the different function hat at least similar types
- [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
- [ ] 316: equation extends into margin
- [ ] 331: I think the application of $w_c$ is wrong (same with $w_r$ and $w_i$ later), shouldn't the : be a comma and what is q in the middle?
- [ ] 336: how are states (which are in $Q$) pushed onto the stack if $\Gamma = \Delta_c \times P$?
- [ ] 339: what are $j$ and $c$, do you mean $i$ and $d$?
- [ ] 341: shouldn't there be a transition for returning from a measure?
- [ ] 348: for?  do you mean accepted by?  but the language of an swA isn't defined either
- [ ] 354: readers may have an easier time seeing this argument if you say that $S$ is extremal.  You should say in the preliminaries that S is extremal if and only if $\leq_\oplus$ is total.
- [ ] 371: How is it ensured that the symbol $<c,p>$ isn't popped and pushed again later during the computation?
- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
- [ ] 379: Almost all update equations need to compute infinite sums.  Don't you think that the assumption of $\bar\Phi$ being effective is pretty strong and gives your the result `for free'?  I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
- [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
- [ ] 654: an intuition for the $\bar q$ would have made understanding the construction far easier

R3 :

- [ ] l. 100. Shouldn't the first measure end before the second starts? This would make $O$ a forest rather than a tree - easily fixed with a "piece" root element.
- [ ] l. 220. The sentence starting here is somewhat confusing (and possibly wrong, depending on if $q'$ and $q''$ are intended to be distinct). I suppose that the point is that one can quit the recursion "early" whenever $w_{ij}$ returns $0$, but the language could be clearer, something like "Since $0$ is absorbing for $\times$, and the identity element for $+$ in $S$, if $w_{ij}(q,a,b,q'')$ is $0$ (meaning there is no transition from $q$ to $q'$ reading $a$ and $b$) that entire term can be ignored in the sum."
- [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2
- [ ] l. 434: $swM$??

List of mistakes:
- [x] 17: who -> which; missing comma after swT
- [x] 18: superfluous comma
- [x] 26: compilers or interpreters for programming languages
- [x] 27: superfluous comma
- [x] 32: superfluous comma
- [x] 33: Western
- [x] 43: skip commas around parentheses; `with' is not an optimal conjuction 
- [x] Figure 1: do no start sentences with symbols; add an `and' before $S$
- [x] 47: Boolean Formulas
- [ ] 51--53: the jump from automata to grammars is uneasy and not really necessary, as the paper mainly deals with automata
- [ ] 57: weighted automata don't have ASTs, but runs
- [ ] 60: other approaches use multioperator monoids or valuation monoids, so `at minima a semiring' is wrong
- [x] 70: the second comma is superfluous
- [x] 74: missing period and space after $\oplus$
- [ ] 75: maybe cite Parsing as intersection by Nederhof and Satta (2003)
- [x] 77: missing space after period
- [x] 84: convenient in this context -> for convenience?
- [x] 87: superfluous comma
- [ ] 116: $\oplus$ `combines two values into a single one' as well (in fact, every   binary operation does)
- [x] 118--119: I would say: … $\leq_\oplus$ which is called the natural ordering of $S$ and is, for every …, defined by …
- [x] Also, you should be consistent with using  'for all' or `for every' in quantifications (I would prefer for every)
- [x] 122: if it IS; I'd also add a comma after total
- [x] 123: Lemma 2 is Lemma 2 in [21], you should refer to it
- [x] 125: you write `then', but the following sentence seems to be disconnected from the previous lemma
- [ ] 126: you should cite Knuth's generalization of Dijkstra's algorithm for superiority, he defines the notion for functions of arbitrary arity, so not just semirings (but multioperator monoids); you should give an example for superiority in the semiring notation (with $\otimes$)
  Moreover, the `corresponds' is wrong, since superiority generalizes non-negative edge weights
- [ ] PS: Knuth's definition of superiority includes monotonicity

@@

- [ ] 299: every line expands into the margin; moreover, in the rightmost equation of line three, in the application of $w_c$, $p'$ and $q'$ are swapped
- [x] 310: topupmost -> topmost
- [ ] 316: equation extends into margin
- [x] 330--331: as indicated by the stack top: better explain that stack symbols are time intervals

@@

- [x] 375: converge (singular)
- [ ] 376: I do not understand the and at termination, and its correctness part.  What do you want to express?
- [ ] 379: Almost all update equations need to compute infinite sums.  Don't you think that the assumption of $\bar\Phi$ being effective is pretty strong and gives your the result `for free'?  I think you should at least elaborate on this and say to what degree your effective label theory is a generalization of the decidable label theory of unweighted symbolic automata.
- [x] 386: We assume the following input to be given
- [x] 394: the notation "sw-parsing" is undefined
- [x] 416: word order: put "to $T$" in the end
- [x] 420: "the sw-parsing problem" or "sw-parsing", but not "the sw-parsing"
- [x] 424: the triple use and does not work well
- [x] 431: double the
- [x] 435: there is something missing between "here" and "offline"

@@

- [ ] 629: typical references for regular tree grammars are Brainerd (1969) and  Gecseg and Steinby (1984)
- [x] 647: inn -> in
- [x] 650: missing period after function
- [x] 652: I think $\bar w$ should be $\bar w'$

@@

----------- Overall evaluation -----------
Transducers and visibly pushdown automata are used for parsing input words into abstract syntax trees. When the input words carry information from an infinite domain, such as time stamps, the transducers and automata need to be extended to handle these, which is the topic of this paper. The transducers are extended in one direction to handle inputs from the infinite domain, using the concepts from symbolic automata. Transducers are extended in another direction to give multiple outputs with associated weights. These can be used to produce multiple options for intermediate representations before computing the abstract syntax trees. The multiple intermediate representations can be compared and the one with the minimum weight can be picked up for further processing by visibly pushdown automata to construct the abstract syntax tree. The visibly pushdown automata are again extended to handle symbolic inputs and ouptut weights for different possible runs, which represent different!
  possible abstract syntax trees.

The contributions of this paper are extending transducers and automata to handle symbolic inputs and to output weights in a semiring instead of Boolean values. The paper also gives procedures for computing minimum weights for given inputs and to compute witnesses for these minimum weights, which will be helpful to construct syntax trees. The definitions have been extended from the respective symbolic and weighted models. The technical details are mostly well written and complemented with intuitive explanations using a running example. However, for now, I am still giving weak accept instead of accept, since I am having difficulty understading a few points. See the last two comments below.

- [ ] Line 209:  "T is based on an intermediate function ...": should this be "The semantics of $T$ is based on an intermediate function ..."?

- [x] Line 281: "artitioned into three" -> "partitioned into three"

@@

- [x] l. 276: generalizing themselves -> themselves generalizing

- [x] l. 281: artitioned -> partitioned

- [ ] l. 288: same as on line 100, measure 1 should maybe end before measure 2

- [x] l. 290: ornemental -> ornamental

- [x] l. 306: reads and pop from stack -> reads and pops from the stack

- [x] l. 307: I would use "computations" or something like that rather than "transitions" here

- [x] l. 310: topupmost -> topmost, though I'm not sure if it's necessary to describe the basic workings of a stack. If you choose to do so, then also

- [x] l. 311: content -> pair

- [x] l. 323: particular -> special

- [x] l. 365: "case of in the" -> either "case in the" or "case of the"

- [x] l. 366: pop -> popped

 - [x] The second case in Figure 3 should start with $d_\top$, not $d_\bot$

- [x] l. 395: I would omit "output". If you want to draw attention to the different alphabets involved, I would say "between the string $s$ over the input alphabet, and the weighted language over the output alphabet defined by $A$"

- [x] l. 399: correspondance -> correspondence

- [x] l. 402: "weight a second time division" -> something like "weight assigned to an additional second time division"

- [x] l. 416: extra commen between "$T$" and "of new transition"

- [x] l. 417: Corolary -> Corollary

- [x] l. 420: This whole paragraph should be shaped up, writing wise.

### 2021-09-01 app-trees.tex
passe section 4 swVPA
corrected review typos

The hierarchical structure of nested-words, defined with the \emph{call} 
and \emph{return} markup symbols
suggest a correspondence with trees.
The lifting of this correspondence to languages, of tree automata and VPA,
has been discussed in~\cite{AlurMadhusudan09nested},
and~\cite{Caralp12VPAmult} for the weighted case.
In this section, we describe a correspondence between the symbolic-weighted extensions
of tree automata and VPA.

@@

\noindent
We associate to $\hat\Omega$
a label theory $\hat{\Phi}$
like in Section~\ref{sec:SWVPA-def},
%
\noindent
and we define a linearization of trees of $\T_\Omega$ into
words of ${\hat\Omega}^*$ as follows:
\begin{description}
\item $\lin(a) = a$ for all $a \in \Omega_0$,
\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) =
       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$
       when $b \in \Omega_k$ for $1 \leq k \leq M$.
\end{description}
%
\begin{example}
The trees in Figure~\ref{fig:score-tree}
represent the two scores in Examples~\ref{ex:running},\ref{ex:nested-word}, 
and their linearization are respectively~$O$ and~$O'$ inn the same examples.
\end{example}

\begin{proposition}\label{lem:SWTA}
For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
there exists an effectively constructible \SWVPA $A'$ over
$\hat\Omega$, $\Semiring$ and $\hat\Phi$
such that for all $t \in \T_\Omega$, $A'\bigl(\lin(t)\bigr) = A(t)$.
\end{proposition}
%
\begin{proof}
Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function
    %
We build
$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
%computing over $\Delta = \Omegai \uplus \Omegac \uplus \Omegar$,
where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$,
of length at most $M$, including the empty sequence denoted by~$\varepsilon$,
and where $P' = Q'$ and $\bar\wei$ is defined by:

>>>>

The hierarchical structure of nested words, defined with the \emph{call} 
and \emph{return} markup symbols
suggest a correspondence with trees.
The lifting of this correspondence to languages, of tree automata and VPA,
has been discussed in~\cite{AlurMadhusudan09nested},
and~\cite{Caralp12VPAmult} for the weighted case.
In this section, we describe a correspondence between the symbolic-weighted extensions
of tree automata and VPA.

@@

\noindent
We associate to $\hat\Omega$
a label theory $\hat{\Phi}$
like in Section~\ref{sec:SWVPA-def},
%
\noindent
and we define a linearization of trees of $\T_\Omega$ into
words of ${\hat\Omega}^*$ as follows:
\begin{description}
\item $\lin(a) = a$ for all $a \in \Omega_0$,
\item $\lin\bigl( b(t_1, \ldots, t_k)\bigr) =
       \call{b} \; \lin(t_1) \ldots \lin(t_k) \; \return{b}$
       when $b \in \Omega_k$ for $1 \leq k \leq M$.
\end{description}
%
\begin{example}
The trees in Figure~\ref{fig:score-tree}
represent the two scores in Examples~\ref{ex:running},\ref{ex:nested-word}, 
and their linearization are respectively~$O$ and~$O'$ in the same examples.
\end{example}

\begin{proposition}\label{lem:SWTA}
For all \SWTA $A$ over~$\Omega$, $\Semiring$ commutative, and $\bar\Phi$,
there exists an effectively constructible \SWVPA $A'$ over
$\hat\Omega$, $\Semiring$ and $\hat\Phi$
such that for all $t \in \T_\Omega$, $A'\bigl(\lin(t)\bigr) = A(t)$.
\end{proposition}
%
\begin{proof}
Let $A = \< Q, \init, \bar{\wei} >$ where $\bar{\wei}$ is presented as above by a function.
    %
We build
$A' = \< Q', P', \init', \bar{\wei}', \final' >$,
%computing over $\Delta = \Omegai \uplus \Omegac \uplus \Omegar$,
where $Q' = \bigcup_{k=0}^{M} Q^k$ is the set of sequences of state symbols of $A$,
of length at most $M$, including the empty sequence denoted by~$\varepsilon$,
and where $P' = Q'$ and $\bar{\wei}'$ is defined by:
